caDDS: Community Assisted Distributed Database for Sequences

### 2020-06-24 Paper2020ACM/Paper2020.tex
June 24 8am
Revisions June 24 10:30

\subsubsection{CAP theorem for Databases} 
There are 3 principles needed when creating a database. The idea is that only two of the three can be focused on at any given time. Only two since the current technology can only handle two. These principles and limitations need to be taken into consideration when designing the system.

@@

\subsection{NCBI}
The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}. This is the standard genomic database used by many researches and institutions. We will refer this later on as well as the \textit{base system}. This is the website used to access and store most genomic information\cite{campbell}. It uses a client-server architecture. Users may download data via the graphics interface (website), or may be done via file transfer protocols such as FTP or HTTP \cite{biotorrents}. The latter is preferred since it may be done to automate the download of data.

@@

\subsection{Problems of Current Distributed Databases for Genomes}
\begin{enumerate}
    \item No public existing usable public framework on distributed databases for genomes
    \item Existing distributed database system for genomes is not replicable
\end{enumerate}

@@

\begin{itemize}
    \item Design the framework for an information system that speeds up the download and upload of genomic data
    \begin{itemize} 
        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
    \end{itemize}
    \item  Design a distributed database system
    \begin{itemize}
        \item To resolve 3 \& 4 of the Problems of Classical Databases. 
    \end{itemize}
\end{itemize}

@@

\subsubsection{Community Aspect}
The sytem was also described as community assisted. The definition of \textbf{community} to be used here is a \textit{set of people who generally collaborate together to do research}. And the focus is the genome research community. Reemphasizing certain features from the database, that any users can upload data. This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community. Another important community aspect is that users can become data nodes themselves. Meaning these users can increase the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later). Frequently used local data as well will be more easily accessible (since it's on local data nodes, and should have faster access speeds). The concept of closer networks having faster speeds (discussed in the network section) is at play here. Another community aspect will be this paper, publicly describing many aspects of the structure and code. So that this concept may be further developed and even implemented by genome research communities.

@@

%HIII NEED HELP HEREE
\section{Proof of Correctness}
correct if
GET - gets data
POST - puts data
PUT - edits data
DELTE - deletes data

speed should be X

Data aross Data Nodes is consistent

@@

\subsection{Base Method}
In the base method, represented by NCBI. We will abstract the transfer into this algorithm. Assuming as well, constant speed, equal division of bandwidth, all users are getting the same data all at once, and there is a constant overhead before downloading can occur.

@@

\subsubsection{General Case of the Proposed Method for Multiple Data Nodes} 
This is different from the General Case of the Proposed Method for Multiple Data Nodes by a factor of D. N total users and D data nodes will receive data in the speed of DX/N bytes per second with H seconds of an initial overhead. N is the number of users sharing the bandwidth. Downloading data (size of X bytes)  will take DX/(NX) + H or (D/N) + H seconds to transfer. A notable aspect of this general case is that as the number of data nodes increases, the user's speed also increases. Even if the previous feature from the general case of the base method is still at play, (an increase in the number of users decreases the user download speed).

@@

\subsection{Sumamry of the Theoretical Analysis of Speed}
Generally as the number of users increases, the data transfer speed decreases by a constant factor for all methods. There is also a concept of bottleneck when the number of users is much greater than the data speed put out by all methods.

@@

\section{Comparison of base method vs distributed method}

base method (NCBI)
- houses a lot of biological data
- go to website
- one website
- one data source
- single point of failure \cite{seqtorr}
- implemented

proposed method
- ability to scale
- multiple (scalable data nodes) data sources
- multiple (scalable data nodes) data links
- not yet implemented
- can be done / implemented on a local scale (limit nodes to the needs)
- only have data for a certain data community
- only have users from a certain data community
- able to design the APIs for speed, security, consistency, tolerance, etc. depending on community needs

- the section on theoretical speed increase in data speed transfer (data nodes > 1)
- can reduce the data stored to house only locally used and needed data

\section{Recommendations}

Implement this. Describe the technologies used and why. And make the code public. Better yet create a template code to easily copy and implement own data centers.

Implementation problems. The network API will face problems in port access especially in UPD internet. Only 2 ports are currently available.

Have a set of testing APIs and have testing APIs embedded in each component. And the testing consistently checks which data have integrity.

Have security. The API structure is one way of having security in the future even if it isnt implemented yet. Restructure and redesigning the APIs are one way to achieve security. Adding user authentication is one way to achieve this. As well as having a separate layer between user and master. Where all interactions happen in that layer (even data transfer??). 
Abstract this project even further. Creating means for master nodes to communicate and collaborate with other master nodes. This will let different caDDS (assuming this is implemented) to communicate and collaborate (send data) to other caDDS.

Abstract this project for non-genomic data as well. To develop the idea of a distributed data community (where distributed databases are easy to scale, implement, and connect). Which can help share many kinds of scientific data more easily. A specific suggestion is for other fields with large data gathered, such as space image data.

>>>>

\subsubsection{CAP theorem for Databases} \label{cap}
There are 3 principles needed when creating a database. The idea is that only two of the three can be focused on at any given time. Only two since the current technology can only handle two. These principles and limitations need to be taken into consideration when designing the system.

@@

\subsection{NCBI} \label{NCBI}
The National Center for Biotechnology Information or more commonly known as NCBI, which discussed briefly in the section on \textit{BioTorrents} and in the \textit{Introduction}. This is the standard genomic database used by many researches and institutions. We will refer this later on as well as the \textit{base system}. This is the website used to access and store most genomic information\cite{campbell}. It uses a client-server architecture. Users may download data via the graphics interface (website), or may be done via file transfer protocols such as FTP or HTTP \cite{biotorrents}. The latter is preferred since it may be done to automate the download of data.

@@

\subsection{Problems of Current Distributed Databases for Genomes}
\begin{enumerate}
    \item No existing usable public framework on distributed databases for genomes
    \item Existing distributed database system for genomes is not replicable
\end{enumerate}

@@

\begin{itemize}
    \item Design the framework for an information system that speeds up the download and upload of genomic data
    \begin{itemize} 
        \item To resolve 1 \& 2 of the Problems of Classical Databases. 
    \end{itemize}
    \item  Design a distributed database system
    \begin{itemize}
        \item To resolve 3 \& 4 of the Problems of Classical Databases, and 1 \& 2 of the Problems of Distributed Databases. 
    \end{itemize}
\end{itemize}

@@

\subsubsection{Community Aspect} \label{community}
The sytem was also described as community assisted. The definition of \textbf{community} to be used here is a \textit{set of people who generally collaborate together to do research}. And the focus is the genome research community. Reemphasizing certain features from the database, that any users can upload data. This means all users in the community can contribute to the knowledge base, and contribute data to be used by the community. Another important community aspect is that users can become data nodes themselves. Meaning these users can increase the overall storage of the database, and at the same time theoretically increase data speeds (which will be discussed more later). Frequently used local data as well will be more easily accessible (since it's on local data nodes, and should have faster access speeds). The concept of closer networks having faster speeds (discussed in the network section) is at play here. Another community aspect will be this paper, publicly describing many aspects of the structure and code. So that this concept may be further developed and even implemented by genome research communities.

@@

\section{Proof of Correctness}
Hii i need help fixing this section di ko pa alam ano ilalagay

The distributed database will function correctly if:

\begin{enumerate}
    \item GET transfers the data to the user with no error
    \item POST transfers the data to the data node with no error
    \item PUT edits the data
    \item DELETE removes the data
    \item all DATA NODES with DATA X have the exact copy of DATA X
\end{enumerate}

@@

\subsection{Base Method}
In the base method, represented by NCBI (see Section \ref{NCBI}). We will abstract the transfer into this algorithm. Assuming as well, constant speed, equal division of bandwidth, all users are getting the same data all at once, and there is a constant overhead before downloading can occur.

@@

\subsubsection{General Case of the Proposed Method for Multiple Data Nodes} 
This is different from the General Case of the Proposed Method for a Single Data Node by a factor of D. N total users and D data nodes will receive data in the speed of DX/N bytes per second with H seconds of an initial overhead. N is the number of users sharing the bandwidth. Downloading data (size of X bytes)  will take DX/(NX) + H or (D/N) + H seconds to transfer. A notable aspect of this general case is that as the number of data nodes increases, the user's speed also increases. Even if the previous feature from the general case of the base method is still at play, (an increase in the number of users decreases the user download speed).

@@

\subsection{Summary of the Theoretical Analysis of Speed} \label{theoreticalspeed}
Generally as the number of users increases, the data transfer speed decreases by a constant factor for all methods. There is also a concept of bottleneck when the number of users is much greater than the data speed put out by all methods.

@@

\section{Comparison of base method vs proposed method}

For the final comparisons of the base method the proposed method, the current base method (NCBI in Section \ref{NCBI}) is the go-to website for biological data and genomic data. It is currently used by and is one website and one server. This points to a single point of failure\cite{seqtorr}. The section \ref{theoreticalspeed} on theoretical analysis of speed described the problem of having a bottleneck when the number of users are too many for the data bandwidth of the server.

The proposed method introduces the idea of data nodes and in section \ref{theoreticalspeed} explains how having at least one data node should theoretically rival the speed of the base method. And having multiple data nodes will be faster in theory than the base method. Another advantage of having a distributed database is the community aspect (see Section \ref{community}). The may be implemented on a local scale, meaning it can have the necessary data for the community. The most frequently used data, may be put on more data nodes increasing the speed of the data transfer (see Section \ref{theoreticalspeed}). Another is that the features on the designs of the speed, consistency, and fault tolerance (see Section \ref{cap} ) may be calibrated to the needs of the community.

\section{Recommendations}
A recommendation is to implement this proposed system. It is advised to have a public code with an easy-to-start documentation.

A small but important technical step is to pay attention to port access especially with university internet, some universities only limit the usable ports to two and this would be an important hurdle to overcome in managing multiple data nodes.

Another is to look into security, which aspects of the code or data must be private or moderated. How does data privacy work? And how does security affect the accessibility of the system? What are the different risks in distributed databases and how to prevent them? Many questions arise in security but these will only be answered better once implementation starts.

A recommendation is to look into a way for master nodes to communicate with other master nodes, this will start the idea of communication and collaboration between caDDS to other caDDS.

The last recommendation is to abstract this project for non-genomic data as well. This is good especially for other scientific communities with its own growing data set such as cosmology with its space image data.
